False Positive Probabilities for all Kepler Objects of Interest: \\ \nvalnew\ newly validated planets and \nfpnew\ likely false positives

### 2016-03-24 document/ms.tex
some updated figures, text, etc.

\newcommand{\ntotal}{7470}
\newcommand{\nfail}{422}
\newcommand{\ncalc}{7048}

\newcommand{\nfp}{424}

\newcommand{\nbadphot}{28} %EmptyPhotometry or BadPhotometry of MissingKOI or IOError
%\newcommand{\nbadphotFP}{34}
%\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{233}
\newcommand{\nbadstellar}{74} %MissingStellar
\newcommand{\nbadsec}{8} %NoWeakSecondary
%\newcommand{\nattempted}{8141}
%\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{38} %MCMCError or NoFitError
\newcommand{\nbadroche}{37}

\newcommand{\nlongfp}{530}

\newcommand{\nreliable}{2823} %from figure
\newcommand{\nval}{1997} %from figure
\newcommand{\nreliableFP}{398} %from figure
\newcommand{\nvalnew}{1325} %from pie chart
\newcommand{\nfpnew}{424}  % all candidate with FPP < 0.9
\newcommand{\ncurval}{984} % from pie chart
\newcommand{\nhz}{10}

@@

In order to properly interpret these results, it is necessary to
understand the range of applicability of the \vespa\ calculation.
First of all, this method selects between different specific explanations 
for the transit-like signal, and cannot
comment on whether the signal might be caused by stellar variability
or an instrumental false alarm.  Thus, \vespa\ results on low
signal-to-noise ratio (SNR) candidates that are not clearly
transit-like must be viewed with caution.  This being said, the reason 
for including the artificial 
``boxy'' and ``long'' models in the model selection calculation
 (\sect{methods:fpp}) is to flag signals that do not 
fit well with any of the astrophysical eclipse models---in fact, the ``long''
model is preferred by \nlongfp\ KOIs that are already dispositioned
as FALSE POSITIVE.

@@

To enable interpretation, \tab{fpp} thus contains the current KOI
disposition, MES, and the results of the positional probability calculations
of \citet{Bryson:KSCI}.  \Fig{fppall} shows FPPs for all KOIs passing
the following criteria:
\begin{itemize}
\item Dispositioned CANDIDATE or CONFIRMED at the NExScI archive,
\item MES $>$ 10, indicating the signal is unlikely to be caused by 
      systematic noise in the light curve,
\item Probability $> 0.99$ to be on the target star, according
  to \citet{Bryson:KSCI}, along with a positional probability
  ``score'' $> \posprobthresh$ (indicating a reliable result), and   
\item $L_{\rm tot} > 10^{-3}$, indicating that not all the hypotheses are
  poor fits to the transit signal.
\end{itemize}

These selections leave \nreliable\ KOIs for which the \vespa\ results
can be considered most reliable;  \Tab{meanFPP} presents the mean FPPs
for different subsets, showing several notable features.  Single KOIs
are about 3$\times$ more likely to be false positives than KOIs in
multiple-KOI systems, in qualitative agreement with
\citet{Lissauer:2012}---and this is true even without giving any
``multiplicity boost'' to multi-KOI systems for the increased transit
probability of subsequent planets once one planet transits in a
coplanar system.  Also, large candidates typically have high FPPs, in
agreement with \citet{Santerne:2012} and \citet{Santerne:2015}.  And
finally, the mean FPPs are very consistent with the \textit{a priori}
predictions of \citet{Morton:2011b} and with the analysis of
\citet{Fressin:2013}.

@@

While the FPP below which to claim planet validation is clearly
an arbitrary choice, there is precedent to using ${\rm FPP} <0.01$ as 
the threshold---\citet{Rowe:2014} validated over 800 multi-planet KOIs
using this number.  Adopting this same threshold, we find that 
of the KOIs with reliable \vespa\ FPPs, \nval\ have ${\rm FPP}
<0.01$, and are thus validated at the 99\% level.  
These KOIs are labeled as such in \tab{fpp}.  These are not all new validations,
however, since a number of them are already CONFIRMED. \Fig{fpppie}
shows a different kind of summary, grouping KOIs by disposition and
splitting up the CANDIDATES according to the \vespa\ results, showing
that \nvalnew\ KOIs are newly validated at the 99\% level.
\Fig{rpcand} shows the radii and periods of the CANDIDATE KOIs,
grouped by FPP.

@@

As there is significant interest in identifying potentially habitable
planets, \tab{hz} lists the properties of \nhz\ CANDIDATE KOIs newly
validated by this work that \textit{may} fall within the optimistic
habitable zones of their host stars \citep{Kopparapu:2013}, 
according to the stellar
properties reported by the DR24 table at the NExScI Archive.  We note
that while more detailed follow-up observations (imaging and
spectroscopy) have been taken for each of these targets, we make no
attempt here to characterize these systems in detail.  Analysis of
high-resolution spectroscopy in particular will solidify the
properties of these host stars, which will affect the habitable zone
boundaries.  We thus emphasize that this list is neither complete nor
final, serving only to draw attention to new validations of 
interest rather than to be a definitive statement on potential 
habitability.  We also note that based solely on the properties in the 
DR24 table, KOI-5475.01 (listed as having a 448-day orbital period) 
should also be included in this list of potentially habitable-zone
planets.  However, as explained in \citet[][Section 5.5.4]{Coughlin:2015},
this particular KOI actually has a 224-day period, making its insolation
too high to be within the habitable zone; we have thus excluded it 
from \tab{hz}.

@@

One potential concern for the reliability of the validations presented
here might be the possibility of unidentified ephemeris matches, or
``period-epoch matches'' (PEMs).  \citet{Coughlin:2014} identified a
number of false positives among KOIs using this method, by comparing
the periods and epochs of KOIs both to each other and to known
catalogs of eclipsing binaries in order to see if any have identical
ephemerides, an indication of contamination, sometimes from far away 
on the \kepler\ CCD.

In the Q1-Q17 DR24 KOI table used as the basis for this work, 980
KOIs were identified as PEMs.  Of these, 187 were not identified by
any other method.  Only 18 of these 187 survived all the quality cuts
described in \sect{results:fpp}. And of these 18, only 4 have FPP 
$< 0.01$.

Thus, we expect only about 0.4\% (4/980) of unidentified PEMs to end
up as validated according to \vespa. As the total rate of PEMs is
estimated to be something around 35\% \citep{Coughlin:2014}, there
remains a small ($\sim$0.1\%  chance) that any of the validated
planets may turn out to be due to  as-yet unidentified contamination
from one of the false positive  sources that cause PEMs.  In other
words, among our validated sample of planets, there may well be one or
two that turn out to be false positives due to distant contamination
(such as via the ``column anomaly'' effect), but not substantially
more.  (See \sect{gaidos} for an example of a KOI caused by a 
column anomaly that went unidentified by the \kepler\ pipeline but
which \vespa\ assigned a high FPP.)

@@

We also briefly discuss another case of individual interest.
KOI-6705.01, a 0.99\,d signal around a mid M-dwarf star, was
identified  as a KOI of possible interest by \citet{Gaidos:2015}.
After significant follow-up observations and detailed full-frame image
analysis, they concluded that the signal was most likely due to a
charge-transfer  effect from a 1.99\,d EB located on the same CCD
column.  For this KOI, \vespa\ calculates an FPP of 1, with the 
``long'' model preferred by far.

probabilities of 0.86 to be a BEB, 0.13 to be a BEB with twice the
period (the true scenario uncovered by the detective work of
\citet{Gaidos:2015}), and 0.01 to be an HEB.  This is an excellent
example of what was discussed in \sect{PEMs}: that even effects like
column anomalies that were not identified by the \kepler\ team as
ephemeris matches will typically be identified as false positives by 
\vespa.

>>>>

\newcommand{\ntotal}{7470}
\newcommand{\nfail}{427}
\newcommand{\ncalc}{7056}

\newcommand{\nbadphot}{28} %EmptyPhotometry or BadPhotometry of MissingKOI or IOError
%\newcommand{\nbadphotFP}{34}
%\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{233}
\newcommand{\nbadstellar}{74} %MissingStellar
\newcommand{\nbadsec}{8} %NoWeakSecondary
%\newcommand{\nattempted}{8141}
%\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{38} %MCMCError or NoFitError
\newcommand{\nbadroche}{39}

\newcommand{\nlongfp}{526}

\newcommand{\nreliable}{2857} %from figure
\newcommand{\nval}{1948} %from figure
\newcommand{\nreliableFP}{419} %from figure
\newcommand{\nvalnew}{1284} %from pie chart
\newcommand{\nfpnew}{428}  % all candidate with FPP < 0.9
\newcommand{\ncurval}{984} % from pie chart
\newcommand{\nhz}{10}

@@

In order to properly interpret these results, it is necessary to
understand the range of applicability of the \vespa\ calculation.
First of all, this method selects between different specific explanations 
for the transit-like signal, and cannot
comment on whether the signal might be caused by stellar variability
or an instrumental false alarm.  Thus, \vespa\ results on low
signal-to-noise ratio (SNR) candidates that are not clearly
transit-like must be viewed with caution.  This being said, the reason 
for including the artificial 
``boxy'' and ``long'' models in the model selection calculation
 (\sect{methods:fpp}) is to flag signals that do not 
fit well with any of the astrophysical eclipse models---in fact, the ``long''
model is preferred ($>50\5$) by \nlongfp\ KOIs that are already dispositioned
as FALSE POSITIVE.

@@

To enable interpretation, \tab{fpp} thus contains the current KOI
disposition, MES, and the results of the positional probability calculations
of \citet{Bryson:KSCI}.  \Fig{fppall} shows FPPs for all KOIs passing
the following criteria:
\begin{itemize}
\item Dispositioned CANDIDATE or CONFIRMED at the NExScI archive,
\item MES $>$ 10, indicating the signal is unlikely to be caused by 
      systematic noise in the light curve, and
\item Probability $> 0.99$ to be on the target star, according
  to \citet{Bryson:KSCI}, along with a positional probability
  ``score'' $> \posprobthresh$ (indicating a reliable result).  
\end{itemize}

These selections leave \nreliable\ KOIs for which the \vespa\ results
can be considered most reliable;  \Tab{meanFPP} presents the mean FPPs
for different subsets, showing several notable features.  Single KOIs
are about 2.5$\times$ more likely to be false positives than KOIs in
multiple-KOI systems, in qualitative agreement with
\citet{Lissauer:2012}---and this is true even without giving any
``multiplicity boost'' to multi-KOI systems for the increased transit
probability of subsequent planets once one planet transits in a
coplanar system.  Also, large candidates typically have high FPPs, in
agreement with \citet{Santerne:2012} and \citet{Santerne:2015}.  And
finally, the mean FPPs are very consistent with the \textit{a priori}
predictions of \citet{Morton:2011b} and with the analysis of
\citet{Fressin:2013}.

One potential concern worth addressing in some detail, before deciding
which planets to validate based on the \vespa\ calculations, is the 
possibility of false positives caused by distant contamination
but not identified through the ``period-epoch match'' (PEM) technique
used by \citet{Coughlin:2014}, due to the fact that not all stars in the 
\kepler\ field were monitored by the mission.  We thus estimate here
the probability that a CANDIDATE KOI to which \vespa\ assigns a low FPP
might still be caused by such a scenario.

In the Q1-Q17 DR24 KOI table used as the basis for this work, 980
KOIs were identified as PEMs.  Of these, 187 were not identified 
as false positives by any other method.  Only 15 of these 187 survived 
all the quality cuts described in \sect{results:fpp}. And of these 15, 
only 3 have FPP $< 0.01$  (See \sect{gaidos} for an example of a KOI caused by a 
``column anomaly'' effect that went unidentified by the \kepler\ pipeline but
to which \vespa\ assigned a high FPP.).  Thus, we expect only about 0.3\% (3/980) of 
as-yet unidentified distant-contamination FPs to end
up with FPP $< 0.01$ according to \vespa. As the fraction of false positives
from pixel contamination but unidentified as PEMs among the entire KOI sample is
estimated to be something around 23\% \citep{Coughlin:2014}, we estimate
that there remains a small ($\sim$0.06\%) residual FPP for all KOIs, even 
when the \vespa\-calculated FPP is negligibly tiny.

@@

While the FPP below which to claim planet validation is clearly
an arbitrary choice, there is precedent to using ${\rm FPP} <0.01$ as 
the threshold---\citet{Rowe:2014} validated over 800 multi-planet KOIs
using this number.  Adopting this same threshold, and incorporating
the 0.06\% residual FPP estimated in \sect{PEMs}, we find that 
of the KOIs with reliable \vespa\ FPPs, \nval\ have ${\rm FPP}
<0.0094$, and are thus validated at the 99\% level.  
These KOIs are labeled as such in \tab{fpp}.  These are not all new validations,
however, since a number of them are already CONFIRMED. \Fig{fpppie}
shows a different kind of summary, grouping KOIs by disposition and
splitting up the CANDIDATES according to the \vespa\ results, showing
that \nvalnew\ KOIs are newly validated at the 99\% level.
\Fig{rpcand} shows the radii and periods of the CONFIRMED and 
CANDIDATE KOIs, with the transparency of the points representing
the \vespa-calculated FPP.

@@

As there is significant interest in identifying potentially habitable
planets, \tab{hz} lists the properties of \nhz\ CANDIDATE KOIs newly
validated by this work that \textit{may} fall within the optimistic
habitable zones of their host stars \citep{Kopparapu:2013}, 
according to the stellar
properties reported by the DR24 table at the NExScI Archive.  We note
that while more detailed follow-up observations (imaging and
spectroscopy) have been taken for each of these targets, we make no
attempt here to characterize these systems in detail.  However, 
we do note that high-resolution imaging of KOI-2418 and KOI-3010 reveal
that these two host stars have close companions that may or may not
change the habitable nature of the planets (e.g., if the planets happen
to transit the secondary star instead of the primary).  Additionally, analysis of
high-resolution spectroscopy will solidify the
properties of all these host stars, affecting the habitable zone
boundaries.  We thus emphasize that this list is neither complete nor
final, serving only to draw attention to new validations of 
interest rather than to be a definitive statement on potential 
habitability.  We also note that based solely on the properties in the 
DR24 table, KOI-5475.01 (listed as having a 448-day orbital period) 
should also be included in this list of potentially habitable-zone
planets.  However, as explained in \citet[][Section 5.5.4]{Coughlin:2015},
this particular KOI actually has a 224-day period, making its insolation
too high to be within the habitable zone; we have thus excluded it 
from \tab{hz}.

@@

@@

We also briefly discuss another case of individual interest.
KOI-6705.01, a 0.99\,d signal around a mid M-dwarf star, was
identified  as a KOI of possible interest by \citet{Gaidos:2015}.
After significant follow-up observations and detailed full-frame image
analysis, they concluded that the signal was most likely due to a
charge-transfer  effect from a 1.99\,d EB located on the same CCD
column.  For this KOI, \vespa\ calculates an FPP of 1, with the 
``long'' model preferred by far (of the astrophysical models, the 
double-period BEB---the true scenario---is preferred). This is an excellent
example of what was discussed in \sect{PEMs}: that even effects like
column anomalies that were not identified by the \kepler\ team as
ephemeris matches will typically be identified as false positives by 
\vespa.
